---
title: "Assignment 1 - The tidyverse"
author: "Finn Krueger"
date: "`r format(Sys.time(), '%B %d, %Y | %H:%M:%S | %Z')`"
output:
  html_document:
    code_folding: show
    df_print: paged
    highlight: tango
    number_sections: no
    theme: cosmo
    toc: no
---
  
<style>
div.answer {background-color:#f3f0ff; border-radius: 5px; padding: 20px;}
</style>

```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      eval = TRUE,
                      error = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      comment = NA)
```

<!-- Do not forget to input your Github username in the YAML configuration up there --> 

***

```{r, include = T}
library(tidyr)
library(dplyr)
library(legislatoR)
library(tibble)
library(lubridate)
library(stringi)
library(stringr)
library(ggplot2)
```

<br>

### Getting started with the Comparative Legislators Database

The Comparative Legislators Database (CLD) includes political, sociodemographic, career, online presence, public attention, and visual information for over 45,000 contemporary and historical politicians from ten countries (see the [corresponding article](https://www.cambridge.org/core/journals/british-journal-of-political-science/article/comparative-legislators-database/D28BB58A8B2C08C8593DB741F42C18B2) for more information). It can be accessed via `legislatoR` - an R package that is available on both [CRAN](https://cran.r-project.org/web/packages/legislatoR/index.html) and [GitHub](https://github.com/saschagobel/legislatoR).

Before you start with the following tasks, skim the tutorial to make yourself familiar with the database. You find it [here](https://cran.r-project.org/web/packages/legislatoR/vignettes/legislatoR.html).

For the following tasks, you will work with ONE of the legislatures. The legislature you'll work with depends on your first name:

| Your first name starts with... | Legislature | Code |
|---------|-------|-------|
| A-C | Austrian Nationalrat | `aut` |
| D-F | Canadian House of Commons | `can` |
| G-I | Czech Poslanecka Snemovna | `cze` |
| J-L | Spanish Congreso de los Diputados | `esp` |
| M-O | French Assemblée | `fra` |
| P-R | German Bundestag | `deu` |
| S-U | Irish Dail | `irl` |
| V-X | UK House of Commons | `gbr` |
| Y-Z | US House of Representatives  | `usa_house` |

The following tasks will be based on data from the `core` and `political` tables of the database.

<br>

***

### Task 1 - Descriptives [8 points in total]

a) What's the overall share of male legislators in the entire dataset? [1 point]

```{r}

# with overall share I assume that you mean Canada ? 
canada_house_core <- get_core(legislature = "can")

share_of_male <- canada_house_core %>%
  filter(sex == 'male')%>%
  nrow()%>%
  print()


share_of_male / canada_house_core %>%
  nrow()%>%
  print()
  
  

```

b) How many (both in absolute and relative terms) legislators died in the same place they were born in? [1 point]

```{r}

Birth_equal_death<- canada_house_core%>%
  filter(birthplace == deathplace)%>%
  nrow()%>%
  print()
  
Birth_equal_death / nrow(canada_house_core)%>%
  print()

```

c) Create a new variable `age_at_death` that reports the age at which legislators died. Then, plot the distribution of that variable for the entire dataset. [2 points]

```{r}
summary(canada_house_core$birth)



canada_age_at_death <- 
  canada_house_core %>%
  filter(!is.na(birth))%>%
  filter(!is.na(death))%>%
  mutate(age_at_death = lubridate::interval(birth, death) / years(1))


hist(canada_age_at_death$age_at_death, xlab = 'Age of legislator at death')




```

d) What is the most frequent birthday in your sample (e.g., “January 15")? [2 points]

```{r}

birthdates <- format(canada_house_core, format="%m/%d")



Frequenca <- birthdates %>%
  filter(!is.na(birth)) %>%
  group_by(birth) %>%
  summarise(freq = (n()))%>%
  arrange(desc(freq))%>%
  print()
  


```

e) Generate a table that provides the 5 legislators with the longest names (in terms of number of characters, ignoring whitespace). [2 points]

```{r}

legislator <- canada_house_core %>%
  mutate(longest_name =  str_replace_all(name, " ", ""))  %>%
  mutate(longest_name =  str_length(longest_name))

table_with_five_legistlators <- legislator %>%
  slice_max(longest_name, n = 5 )%>%
  select(name) %>%
  print()


table_with_five_legistlators <-  head(table_with_five_legistlators, n = 5)

print(table_with_five_legistlators)
```

<br>

***

### Task 2 - Exploring trends in the data [8 points in total]

a) Using data from all sessions, provide a plot that shows the share of male legislators by session! [3 points]

```{r}
can_political <- get_political(legislature = "can")
glimpse(can_political)


canada_political_core <- canada_house_core %>% left_join(can_political, by = "pageid")


canada_session <-canada_political_core %>%
  group_by(session) %>% 
  mutate(male = (sex == 'male'))  %>% 
  summarize(male_ratio = mean(male, na.rm = TRUE), 
            session_transformed = ymd(first(session_start)))


plot(canada_session$session_transformed, 
     canada_session$male_ratio, 
     ylab = "Share of male legislators", 
     xlab = "Year",
     main = "Share of male legislators by session")


```

b) Explore another relation of variables in the dataset. Your analysis should (i) use data from both the `core` and the `political` table and (ii) feature data from several sessions. Also, offer a brief description and interpretation of your findings! [5 points]

```{r}
# I use the already joined dataset canada_political_core <- canada_house_core %>% left_join(can_political, by = "pageid")

ethnicitiy_share <- canada_political_core%>% 
  filter(!is.na(ethnicity))%>% 
  group_by(session, ethnicity)%>% 
    summarise(count = n())  %>%
    mutate(share_ethnicity = count / sum(count))
  
ggplot(ethnicitiy_share, aes(session, share_ethnicity, color = ethnicity)) +
  geom_point() +
  labs(title = "Share of ethnic groups per session", 
       x = "legislative period in sessions", 
       y = "share in percent")  
  
  

```

<div class = "answer">

In this task, I have investigated the diversity in ethnicity of legislators from all sessions. Particularly, I was interested to answer to what extent the share of other ethnicity than white has risen, since the first session from 1872 to the last session beginning in 2019. 

In the X axis of the plot shows each legislative period in sessions, whereas on the Y axis displays the share of arab, asian, black, hispanic, native and white legislators for the given session. 

Interestingly, you can see that the white legislators are still the heavily dominant ethnicity with  87.5% in the current and last session 43 beginning from 2019. From the 15th session that has started in 1925 up until the 24 session that started in 1958 you can see that the entire legislators were white.

Overall, the overall trend that we can observe is that the parliament is becoming more diverse. This trend holds for the overall increase in the share of legislators not being white, but also by the increasing diversity of different ethnicities of the legislators. 

</div>

<br>

***



